---
title: "Pancrea_GSE84133"
output: html_document
---
Description: compute and visualize HIPPO for different z threshold; parallelized computation using library snow 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

## Download the latest HIPPO
## Download GSE84133 Data

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE84133

GSE84133 has multiple batches (both human and mouse) of scRNA-seq data of pancrea tissue. You can download from the above URL under the Supplementary file, "GSE84133_RAW.tar". Download and unzip it. 


```{r, cache=TRUE}
#### set up ####
library(Matrix)
library(data.table)
library(SingleCellExperiment)
library(SC3)
library(Seurat)
library(dplyr)
library(ggplot2)
library(HIPPO)
library(mclust)
library(parallel)
library(MASS)
library(snow)
```

extracting data
```{r, cache=TRUE}
gse84133makesce = function(filename){
  X = read.csv(paste0("C:/Users/lfang/Desktop/working directory for R/GSE84133_RAW/",filename,".csv"))
  Xnew = as.matrix(X[, -c(1:3)])
  rownames(Xnew) = X$barcode
  ind = which(duplicated(X$barcode))
  sce = SingleCellExperiment(assays = list(counts = t(Xnew)))
  colData(sce)$phenoid = X$assigned_cluster
  colData(sce)$barcode = X$barcode
  if(length(ind)>0){
    sce = sce[, -ind]
  }
  tablabel = table(colData(sce)$phenoid)
  ind = names(tablabel)[which(tablabel<=3)]
  cellind = which(colData(sce)$phenoid %in% ind)
  sce = sce[, -cellind]
  rm(X, Xnew)
  return(sce)
}
```

```{r, cache=TRUE}
X1 = gse84133makesce("GSM2230757_human1_umifm_counts")
#X2 = gse84133makesce("GSM2230758_human2_umifm_counts")
```
parallel processing framework 
```{r}
# z=vector('list',4)
# z=1:4
# system.time(lapply(z,function(x) Sys.sleep(1)))
# cl<-makeCluster(8,type="SOCK")
# system.time(clusterApply(cl, z,function(x) Sys.sleep(1)))
```

defining functions
```{r}
rand_VS_z<-function(data2, z){#outputs randindex for different z and k 
  
   hippo_z<-function(X, z){
    X=hippo(X,  
          feature_method="zero_inflation",
          clustering_method="kmeans", 
          K=12,
          z_threshold=z)
    X=hippo_dimension_reduction(X, method="tsne")
  }

  ARI<-function(data, k){#ARI outputs adjusted rand index with diff k 
  truelabel = colData(data)$phenoid
  adjustedRandIndex(truelabel,get_hippo(data)$labelmatrix[,k])
    }
  ARI<-Vectorize(ARI, "k") 
  
  data2=hippo_z(data2, z)
  cat("hippo done for z value", z)
  return(ARI(data2, c(2:12)))
}

###
RandIndexDataset<-function(z_thresh){
  return (rand_VS_z(X1, z_thresh))
}

###
```

prepping parallel computing 
```{r}
numCores=detectCores()
cl<-makeCluster(11,type="SOCK")
registerDoParallel(cl)

clusterExport(cl,c("rand_VS_z", "hippo", "X1", "hippo_dimension_reduction", "adjustedRandIndex", "colData", "get_hippo"), envir = .GlobalEnv)

z_threshold=c(0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75)

```

```{r}
set.seed(35)

parLapply(cl, z_threshold, RandIndexDataset)
#if no seed/bad seed (eg setseed36) runs into "下标出界" at times 

#system.time(parLapply(cl, z_threshold, RandIndexDataset)) 
#用户,系统,流逝 0.50,0.65, 676.93 
```

for loop system time 
```{r}
# system.time(
#    for (i in c(1:7)){
#     cat(i)
#      RandIndexDataset(z_threshold[i])
#  })
#2855.97   16.07 3341.14 
```

## Run HIPPO 

"SC3" tries clustering under multiple settings and combines the result to get the consensus result. Be prepared to run the algorithm for a long time, possibly up to an hour, when you use SC3. You need one extra parameter called sc_n_cores, which determines the number of cores in your local machine (laptop) to use. My laptop crashed in the middle when I tried 3, so maybe play with 1 or 2. You can try more if you have a good laptop.

plotting using ggplot2
```{r}
ggplot(data=RZdataframe, aes(x=c(2:12))) +
  geom_line(aes(y=unlist(RZdataframe[1])))+
  geom_line(aes(y=unlist(RZdataframe[2])))+
  geom_line(aes(y=unlist(RZdataframe[3])))+
  geom_line(aes(y=unlist(RZdataframe[4])))+
  geom_line(aes(y=unlist(RZdataframe[5])))+
  geom_line(aes(y=unlist(RZdataframe[6])))+
  geom_line(aes(y=unlist(RZdataframe[7])))
#yuck 
```

plotting using plot function 
```{r, cache=TRUE}
set.seed(35)
#create plot 

#plotting data for first z threshold value
plot(c(1:ncol(rand_VS_z_data)), rand_VS_z_data[,1], type="l", col=1, 
     ylim=c(min(rand_VS_z_data), max(rand_VS_z_data)), 
     main="comparing different z threshold across multiple k values", 
     ylab="Rand Index",
     xlab="number of clusters",
     xaxt="n")

axis(1, at=1:ncol(rand_VS_z_data), lab=c(2:(l+1)))#adding one to x-values 1... n 
#to find corresponding k values 2...n+1 

#adding all other z threshold data
for (i in 2:ncol(rand_VS_z_data)){
lines(c(1:length(diff_z_thres[1,])), diff_z_thres[i,], type="l", col=i)
}

legend(9, 0.5, legend=z_threshold, title="z threshold value",
       col=c(1:7), lty=1, cex=0.8)
#dev.copy(pdf,'myplot.pdf')
#dev.off()

```

Code for visualizing tsne result:

tsnedf1 = get_hippo(X1)$tsne[1:ncol(X1), 1:2]
tsnedf1 = tsnedf1 %>% mutate(truelabel = colData(X1)$phenoid) %>%
  mutate(hippolabel = as.factor(get_hippo(X1)$labelmatrix[,10]))

ggplot(tsnedf1, aes(x = tsne1, 
                   y = tsne2, 
                   col = truelabel, 
                   shape = hippolabel)) +
  geom_point()

```{r}
X2 = hippo(X2, 
           feature_method = "zero_inflation", #play with this parameter
           clustering_method = "SC3", #play with this parameter
           K=3,
           z_threshold = 1) #play with this parameter
```


```{r}
X3 = gse84133makesce("GSM2230759_human3_umifm_counts")
X3 = hippo(X3, ...)
```